An efficient composition of bidirectional programs by tupling and lazy updates

### 2020-02-21 paper/1.introduction.tex
small fix

To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
\textcolor{red}{Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$. However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result. In this case we can use $[..., [3]]$ as a complement.}
The key idea is straightforward: Complements are smaller
% program fragments
than
% the original
intermediate states.
%Therefore this strategy solves the previous two problems.
%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
%
%\vspace{2mm}
%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
%$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
%\vspace{2mm}
%
For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
Let us reconsider our example $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$. Here, two points are worth noting. First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$. Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, this simple
% combined
$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}. To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation. We explain this second method and all optimizations in Section~\ref{sect:xpg}.

>>>>

To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
\textcolor{red}{Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$. However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result of $put$. In this case we can use $[..., [3]]$ as a complement.}
The key idea of the second approach is straightforward: Complements are smaller
% program fragments
than
% the original
intermediate states.
%Therefore this strategy solves the previous two problems.
%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
%
%\vspace{2mm}
%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
%$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
%\vspace{2mm}
%
For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
Let us reconsider our example $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$. Here, two points are worth noting. First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$. Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, this simple
% combined
$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}. To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation. We explain this second method and all optimizations in Section~\ref{sect:xpg}.

### 2020-02-21 paper/2.minbigul.tex
small fix

If we think that a minBiGUL program $bx$ has a type of $\textit{MinBiGUL} \ s \ v$, the type of $bfoldr$ will be look like $\textit{MinBiGUL} \ (a, b) \ a \to \textit{MinBiGUL} \ ([a], b) \ b$. You can easily find the similarity between the above definition of $\textit{bfoldr}$ with the following definition of $\textit{foldr}$:

>>>>

If we think that a minBiGUL program $bx$ has a type of $\textit{MinBiGUL} \ s \ v$, the type of $bfoldr$ will be look like $\textit{MinBiGUL} \ (a, b) \ b \to \textit{MinBiGUL} \ ([a], b) \ b$. You can easily find the similarity between the above definition of $\textit{bfoldr}$ with the following definition of $\textit{foldr}$:

### 2020-02-21 paper/4.xpg.tex
small fix

    A restriction for $\pginline{Case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$. If a tupling occurs, there will be 4 combinations from these conditions. This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied. The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper. This does not happen for the others which is used in the construction of $\pginline{Case}$.
% {\textcolor{red}{If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction. Of course, all BX programs do not always have such efficient way. An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}

@@

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid. Since there is no real view, this $dummy$ is necessary to pair with original source $s$ to form the input of $put \, [\![bx_1]\!]$. In general, $dummy$ depends on the source $s$, the view $v$ and/or the program $bx_1$. Programmers can be required to give a way to construct $dummy$, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is one of the easiest ways to meet our expectation. That setting is used in our experiments. The last equality changes from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively. Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$. The substitution of $v_3$ by $v_3'$ is simply replacing the variable name since $v_3$ and $v_3'$ hold different results of $\getbx{bx_1}{s}$ and $\getbx{bx_1}{s_1}$ respectively. Because both variables are no longer used later, this substitution does not affect the outcome of the function.

>>>>

    A restriction for $\pginline{Case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$. If a tupling occurs, there will be 4 combinations from these conditions. This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied. The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper. Because they evaluate different $bx$ for $put$ and $get$, we can not evaluate them efficiently. This does not happen for the others which is used in the construction of $\pginline{Case}$.
% {\textcolor{red}{If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction. Of course, all BX programs do not always have such efficient way. An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}

@@

The construction of $\pginline{bx_1 \circ bx_2}$ is the most important part in the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid. Since there is no real view, this $dummy$ is necessary to pair with the original source $s$ to form the input of $put \, [\![bx_1]\!]$. In general, $dummy$ depends on the source $s$, the view $v$ and/or the program $bx_1$. Programmers can be required to give a way to construct $dummy$, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is one of the easiest ways to meet our expectation. That setting is used in our experiments. The last equality changes from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively. Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$. The substitution of $v_3$ by $v_3'$ is simply replacing the variable name since $v_3$ and $v_3'$ hold different results of $\getbx{bx_1}{s}$ and $\getbx{bx_1}{s_1}$ respectively. Because both variables are no longer used later, this substitution does not affect the outcome of the function.
